Reproducible Research Peer Assessment 1
==========================================

### Configure initial settings
```{r}
echo = TRUE 
options(scipen = 1) 
library(ggplot2)
```

### Load and process the data

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
data$month <- as.numeric(format(data$date, "%m"))
removeNAdata <- na.omit(data)
rownames(removeNAdata) <- 1:nrow(removeNAdata)
```

### What is the mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.

* Make a histogram of the total number of steps taken each day

```{r}
ggplot(removeNAdata, aes(date, steps)) + geom_bar(stat = "identity", colour = "red", fill = "red", width = 1) + facet_grid(. ~ month, scales = "free") + labs(title = "Total Number of Daily Steps", x = "Date", y = "Number of steps")
```


* Calculate and report the mean and median total number of steps taken per day

Mean total number of steps taken per day:

```{r}
totalSteps <- aggregate(removeNAdata$steps, list(Date = removeNAdata$date), FUN = "sum")$x
mean(totalSteps)
```

Median total number of steps taken per day:
```{r}
median(totalSteps)
```

### What is the average daily activity pattern?
* Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
avgSteps <- aggregate(removeNAdata$steps, list(interval = as.numeric(as.character(removeNAdata$interval))), FUN = "mean")
names(avgSteps)[2] <- "meanOfSteps"

ggplot(avgSteps, aes(interval, meanOfSteps)) + geom_line(color = "red", size = 1) + labs(title = "Plot of Steps Taken in 5-minute Intervals", x = "5-minute intervals", y = "Average Number of Steps Taken")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
avgSteps[avgSteps$meanOfSteps == max(avgSteps$meanOfSteps), ]
```

### Imputing missing values
* The total number of data rows with NAs:

```{r}
sum(is.na(data))
```

* Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

* Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}
tidyData <- data 
for (i in 1:nrow(tidyData)) {
    if (is.na(tidyData$steps[i])) {
        tidyData$steps[i] <- avgSteps[which(tidyData$interval[i] == avgSteps$interval), ]$meanOfSteps
    }
}
sum(is.na(tidyData))
```


* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 


```{r}
ggplot(tidyData, aes(date, steps)) + geom_bar(stat = "identity",
                                             colour = "red",
                                             fill = "red",
                                             width = 1) + facet_grid(. ~ month, scales = "free") + labs(title = "Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```

* Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Mean total number of steps taken per day:

```{r}
newTotalSteps <- aggregate(tidyData$steps, 
                           list(Date = tidyData$date), 
                           FUN = "sum")$x
newMean <- mean(newTotalSteps)
newMean
```

Median total number of steps taken per day:

```{r}
newMedian <- median(newTotalSteps)
newMedian
```

After imputing the missing data, the mean is the same and the median is slightly greater.

### Are there differences in activity patterns between weekdays and weekends?

* Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
tidyData$weekdays <- factor(format(tidyData$date, "%A"))
levels(tidyData$weekdays)
levels(tidyData$weekdays) <- list(weekday = c("Monday", "Tuesday",
                                             "Wednesday", 
                                             "Thursday", "Friday"),
                                 weekend = c("Saturday", "Sunday"))
levels(tidyData$weekdays)
table(tidyData$weekdays)
```

* Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).


```{r}
avgSteps <- aggregate(tidyData$steps, 
                      list(interval = as.numeric(as.character(tidyData$interval)), 
                           weekdays = tidyData$weekdays),
                      FUN = "mean")
names(avgSteps)[3] <- "meanOfSteps"
library(lattice)
xyplot(avgSteps$meanOfSteps ~ avgSteps$interval | avgSteps$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")
```

